% !TeX root = ..\..\notes-main.tex
\documentclass[../notes-main.tex]{subfiles}
\begin{document}
\section{Introduction to Modeling}
\fxerror{Missing content.}
\begin{mdframed}
    \begin{center}
        \redtext{\emph{Common man} observes the physical phenomena around him; the \emph{scientist} encodes them into models.}
    \end{center}
\end{mdframed}
\label{fig:modeling-intro}
\fxerror{Content below is not 1 to 1, needs to be rewritten and formatted.}
\section{Definition of a Model}
\begin{itemize}
    \item A model of a system is a tool we use to answer questions about the system without having to do an experiment.
    \item A collection of mathematical relationships between system/process variables which purports to describe the behavior of a physical system.
    \item A convenient surrogate of the physical system.
\end{itemize}
\section{What is it's main use?}
% Section title was wrong. Still needs a rewrite.
\fxerror{Content below is a mess.}
\begin{itemize}
    \item To investigate system response under various input conditions both rapidly, and inexpensively, without tampering with the actual physical entity.
    \item Analytically design controllers.
\end{itemize}
\fxerror{There's a whole missing section between these two. rather six slides\dots}
\section{Basic Concepts: Poles, Zeros, and System Order}
\fxerror{The following content is not correct, title is misleading, and content does not match the slides.}
\textbf{Poles and Zeros}
\begin{itemize}
    \item Physical phenomena can be categorized essentially into:
    \begin{enumerate}
        \item Energy dissipation
        \item Energy absorption or storage
    \end{enumerate}
    \item Familiar models: Resistance, Capacitance, and Inductance
    \begin{itemize}
        \item Resistance models the phenomenon of energy dissipation.
        \item Capacitance and Inductance model the phenomenon of energy absorption or storage.
    \end{itemize}
\end{itemize}

\textbf{System Order}
\fxerror{add required content, does not match slides.}
\begin{itemize}
    \item The order of a system equals the number of independent energy-storing elements of the system.
    \item Example:
    \begin{itemize}
        \item R-Circuit: No. of Energy storing elements = 0 Order = 0
        \item R-L Circuit: No. of Energy storing elements = 1 Order = 1
        \item R-C Circuit: No. of Energy storing elements = 1 Order = 1
        \item R-L-C Circuit: No. of Energy storing elements = 2 Order = 2
    \end{itemize}
\end{itemize}
\fxerror{Cast chat GPT to the depths of hell. This is not the content I need.}
\end{document}
